

```{r}
library(tidyverse)
```

# Part 1

```{r}
df = read_csv("Cook_County_Assessor_s_Residential_Modeling_Characteristics__Chicago_.csv")
df
```

```{r}
names(df)
```

```{r}
library(stringr)

names(df) = tolower(str_replace_all(pattern = "[ \\t\\s]", replacement = "_", string = names(df)))
names(df) = str_replace_all(string = names(df), pattern = "[^a-z0-9_]+", replacement = "")
names(df) = str_replace_all(string = names(df), pattern = "__", replacement = "_")
names(df)
```


# Part 2

How many single family homes are included in the data? 
Out of all the residential properties, what is the most common class? 

```{r}
df %>% 
  group_by(use) %>%
  count() %>%
  filter(Use == 'Single-Family')
```

```{r}
df %>%
  group_by(property_class) %>%
  count() %>%
  ungroup() %>%
  slice_max(n)
```

# Part 3

Using lubridate, calculate some information on sales:

- Number of sales in any January
- Number of sales in 2020
- Number of sales on January 2020
- Number of sales on your birthday (or favorite day)
- Number of sales on Wednesday (or your favorite day of the week)

```{r}

```

